The ChEMBL database as Linked Open Data

### 2012-07-09 article.tex
Cleaning up the first part of the Results section

\subsection*{ChEMBL-RDF}

\subsubsection*{Data Structure}

For each of the common resource classes, a triple pattern was defined, following the
data available in the relational database. Figure~\ref{f1} shows how the various resource
classes are linked together. This section will show how data in the ChEMBL database
is exposed at an triple level.

@@

The core concept in the ChEMBL database is that of the biological activity. This
is the type of information that the ChEMBL database extracts from literature.
In ChEMBL-RDF activities are mostly represented using the custom ChEMBL-RDF vocabulary, where the
type, fields and links to other resources are using the \textit{chembl}
namespace. While ChEMBL provides both the original activities as found in literature
and standardized values allowing comparison between studies, the triples only
make the latter available. The CiTO ontology is used to point to the paper from
which the data was extracted.

@@

More than five thousand different activity types are captured by the ChEMBL database.
The top five types are Potency (43\%), IC50 (13\%), MIC (4.6\%), Inhibition (3.7\%),
and Ki (3.6\%). The activity types in ChEMBL-RDF are currently not available as, or
using, an ontology.

The activities themselves are measured against assays, which make up a second important
resource type. Various assay types are found in the database: chembl:ADMET, chembl:Binding,
chembl:Functional, chembl:Property, and chembl:Unassigned. The assay information link
activities to a target (see below) and has a short description. The confidence information
assigned by the ChEMBL is exposed too, and was previously used to improve activity
prediction using Bayesian statistics~\cite{Willighagen2011}.

\begin{small}
\begin{verbatim}
assay:a17
 a chembl:Assay ;
 cito:citesAsDataSource
  res:r11347 ;
 chembl:hasAssayType chembl:ADMET ;
 chembl:hasConfScore "7"^^xsd:int ;
 chembl:hasDescription
  "Inhibition of ... hydroxylase" ;
 chembl:hasTarget target:t100122 .
\end{verbatim}
\end{small}

The assays measure activities against particular biological targets. The ChEMBL database
recognizes various types: pro:PR\_0000001, chembl:ADMET, chembl:CELL-LINE,
chembl:NUCLEIC-ACID, chembl:ORGANISM, chembl:SUBCELLULAR, chembl:TISSUE,
chembl:UNCHECKED, and chembl:UNKNOWN. The latter two are currently defined as
explicit types.

@@

\subsubsection{Data Statistics}

\# triples
\# links out
\# problems

The full set of triples are available from the website http://semantics.bigcat.unimaas.nl/chembl12/.

FIXME!

In contrast to earlier releases of ChEMBL-RDF, it now contains the chemical properties
provided by the ChEMBL database as calculated by ACD/Labs and XXXXX. The data includes
chemical properties like polar surface area, pKa and logP, counts for hydrogen bond donor
and acceptor, and rotational bonds. This data is provided for XXXXX structures.

\subsubsection*{Data Statistics and Validation}

>>>>

\subsection*{Data Structure}

For each of the common resource classes, a triple pattern was defined, following the
data available in the relational database. Figure~\ref{f1} shows how the various resource
classes are linked together, detailed in this section at a triple level.

@@

The core concept in the ChEMBL database is that of the biological activity. This
is the type of information that the ChEMBL database extracts from literature.
Standard ontologies are being explored for future release for annotation
of activities, but, currently, the
type, fields and links to other resources are still using the \textit{chembl}
namespace. While ChEMBL provides both the original activities as found in literature
and standardized values allowing comparison between studies, the triples only
make the latter available. The link to literature is encoded using the CiTO ontology.
This results in a set of triples that look like:

@@

More than five thousand different activity types, represented by the chembl:type predicate,
are captured by the ChEMBL database.
The top five types are Potency (43\%), IC50 (13\%), MIC (4.6\%), Inhibition (3.7\%),
and Ki (3.6\%). The activity types in ChEMBL-RDF are currently not available as, or
using, an ontology.

The activities themselves are measured against assays, which make up a second important
resource type. Various assay types are found in the database: chembl:ADMET, chembl:Binding,
chembl:Functional, chembl:Property, and chembl:Unassigned.

\begin{small}
\begin{verbatim}
assay:a17
 a chembl:Assay ;
 cito:citesAsDataSource
  res:r11347 ;
 chembl:hasAssayType chembl:ADMET ;
 chembl:hasDescription
  "Inhibition of ... hydroxylase" .
\end{verbatim}
\end{small}

Each assay measures activities against one or more particular biological targets. To each
target the ChEMBL database associates a confidence score. Because this score is specific
for each target, the following construct is used:

\begin{small}
\begin{verbatim}
assay:a17 chembl:hasTargetScore assay:a17/score/t100122 .
assay:a17/score/t100122
 chembl:forTarget target:t100122 ;
 chembl:hasConfScore "7"^^xsd:int .  
\end{verbatim}
\end{small}

The ChEMBL database
recognizes various types: pro:PR\_0000001, chembl:ADMET, chembl:CELL-LINE,
chembl:NUCLEIC-ACID, chembl:ORGANISM, chembl:SUBCELLULAR, chembl:TISSUE,
chembl:UNCHECKED, and chembl:UNKNOWN. The latter two are currently defined as
explicit types.

@@

\subsection*{Data Statistics and Validation}
